R for Clinical Study Reports and Submission

### 2021-11-05 tlf-efficacy.Rmd
general updates to improve readability

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize primary and secondary efficacy endpoints,
in Section 11.4, Efficacy Results and Tabulations of Individual Participant.

@@

For efficacy analysis, we only analyze the change from baseline glucose data at week 24.

@@

To prepare the analysis dataset,
we need both `adsl` and `adlbc` datasets for this analysis.

@@

We first define the analysis dataset using
efficacy population flag `EFFFL` and
all records post baseline (`AVISITN > 1`) and on or before Week 24 (`AVISITN <= 24`).
Here the variable `AVISITN` is the numerical analysis visit.
For example, if the analysis visit is recorded as "Baseline", i.e., `AVISIT = Baseline`, then `AVISITN = 0`;
if the analysis visit is recorded as "Week 24", i.e., `AVISIT = Week 24`, then `AVISITN = 24`;
if the analysis visit is blank, then `AVISITN` is also blank.
We will discuss these missing values in Section 6.4.

@@

We first summarize observed data at Baseline and Week 24

@@

We also summarize observed change from baseline glucose at Week 24.

@@

We start to analyze the imputed data using the ANCOVA model with treatment and baseline glucose as covariates.

@@

Then we use the `emmeans` R package to obtain
within and between group least square mean (LS mean)

@@

We combine `t11`, `t12` and `t13` to get the first part of the report data

@@

Then we use `r2rtf` to prepare the table format for `t1`.
We also highlight how to handle special characters in this example.

@@

- Step 1: Read the data into R, i.e., `adsl` and `adlb`.
- Step 2: Define the analysis dataset. In this example, we define two analysis datasets. The first dataset is the efficacy population (`gluc`). The second dataset is the collection of all records post baseline and on or before week 24 (`ana`).
- Step 3: Impute the missing values. In our example, we name the `ana` dataset after imputation as `ana_locf`.
- Step 4: Summarize `gluc`, i.e., calculate the mean and standard derivation, and then format it into an RTF table.
- Step 5: Summarize `ana_locf`, i.e., calculate the pairwise comparison by ANCOVA model, and then format it into an RTF table.
- Step 6: Rowly bind the output in Step 4 and Step 5.

>>>>

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
primary and secondary efficacy endpoints need to summarized
in Section 11.4, Efficacy Results and Tabulations of Individual Participant.

@@

For efficacy analysis, only the change from baseline glucose data at week 24 is analyzed.

@@

To prepare the analysis, both `adsl` and `adlbc` datasets are required.

@@

First, both the population and the data in scope are selected.
The analysis is done on the efficacy population, identified by `EFFFL == "Y"`, and
all records post baseline (`AVISITN >= 1`) and on or before Week 24 (`AVISITN <= 24`).
Here the variable `AVISITN` is the numerical analysis visit.
For example, if the analysis visit is recorded as "Baseline", i.e., `AVISIT = Baseline`, then `AVISITN = 0`;
if the analysis visit is recorded as "Week 24", i.e., `AVISIT = Week 24`, then `AVISITN = 24`;
if the analysis visit is blank, then `AVISITN` is also blank.
We will discuss these missing values in Section 6.4.

@@

First the observed data at Baseline and Week 24 is summarized:

@@

Also the observed change from baseline glucose at Week 24 is summarized:

@@

The imputed data is analyzed using the ANCOVA model with treatment and baseline glucose as covariates.

@@

The `emmeans` R package is used to obtain
within and between group least square mean (LS mean)

@@

`t11`, `t12` and `t13` are combined to get the first part of the report data

@@

Then `r2rtf` is used to prepare the table format for `t1`.
We also highlight how to handle special characters in this example.

@@

- Step 1: Read the data into R, i.e., `adsl` and `adlb`.
- Step 2: Define the analysis dataset. In this example, we define two analysis datasets. The first dataset is the efficacy population (`gluc`). The second dataset is the collection of all records post baseline and on or before week 24 (`ana`).
- Step 3: Impute the missing values. In our example, we name the `ana` dataset after imputation as `ana_locf`.
- Step 4: Summarize `gluc`, i.e., calculate the mean and standard derivation, and then format it into an RTF table.
- Step 5: Summarize `ana_locf`, i.e., calculate the pairwise comparison by ANCOVA model, and then format it into an RTF table.
- Step 6: Set the results from Step 4 and Step 5 and output.
